---
title: "MeaSeq: Measles Sequencing Run Report"
output: 
  flexdashboard::flex_dashboard:
    self-contained: true
    orientation: rows
    vertical_layout: scroll
    mathjax: NULL
    theme:
      version: 4
      bootswatch: flatly
    logo: phac.svg
params:
  strain: ""
  overall_qc: ""
  name: ""
  email: ""
  phone: ""
  website: ""
---

<style>

.navbar-logo img {
    position: relative;
    margin-left: 5px;
    height: 30px;
}
.navbar-fxixed-top {
  margin-bottom: 20px;
}

body {
  padding-top: 60px;
}

footer {
  position: fixed;
  bottom: 0;
  width: 100%;
  background-color: #f1f1f1;
  text-align: center;
  padding: 0px;
  font-size: 12px;
  color: #333;
}
</style>

<!-- --------------------------------------- -->
<!-- Setup: libraries, parameters, functions -->
<!-- --------------------------------------- -->
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
# Load in libraries
suppressPackageStartupMessages(library(rmarkdown))
suppressPackageStartupMessages(library(flexdashboard))
suppressPackageStartupMessages(library(plotly))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(htmltools))
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(shiny))
suppressPackageStartupMessages(library(shidashi))

# Assign imported parameters
strain <- params$strain
overall_qc <- params$overall_qc
name <- params$name
email <- params$email
phone <- params$phone
website <- params$website

# Create Header function
catHeader <- function(text = "", level = 3) {
  cat(
    paste0(
      "\n\n", 
      paste(rep("#", level), collapse = ""), 
      " ",
      text,
      "\n\n"
    )
  )
}

# Set report date
rep_date <- Sys.Date()

# Footer
footer_str <- paste(
              "Report generated using MeaSeq on",
              rep_date,
              "- © Government of Canada, Public Health Agency of Canada, National Microbiology Laboratory",
              sep = ' '
            )

# Call a plotly plot so it can load later within for loops
htmltools::tagList(plot_ly(x = rnorm(10), type = "histogram"))

# Horizontal Line function for plotly
hline <- function(y = 0, color = "black") {
  list(
    type = "line",
    x0 = 0,
    x1 = 1,
    xref = "paper",
    y0 = y,
    y1 = y,
    line = list(color = color, dash = "dash")
  )
}
```

<!-- ------------------ -->
<!-- Setup: Sample Data -->
<!-- ------------------ -->
```{r interactive-table-setup}
# Read multiqc data
summary_df <- read_csv(overall_qc, locale = locale(encoding = "UTF-8"))
# Create a vector of sample names
samples <- summary_df$sample

# Add in 0's for specific NA value columns
summary_df <- summary_df %>%
  mutate_at(vars(`genome_completeness_percent`, `mean_sequencing_depth`, `median_sequencing_depth`), ~ ifelse(is.na(.), 0, .))

# Set values for value boxes
total_samples_count <- nrow(summary_df)
failed_samples_count <- sum(summary_df[["genome_completeness_percent"]] < 50)
passing_samples_count <- total_samples_count - failed_samples_count

# Create vector of failed samples
failed_samples <- summary_df %>%
  filter(`genome_completeness_percent` < 50) %>%
  pull(sample)

# Create formatting conditions for Warnings
summary_df[summary_df == ""] <- NA
for (i in 1:nrow(summary_df)) {
  ifelse(
      summary_df[[i,"strain"]] != strain |
      is.na(summary_df[[i,"strain"]]) |
      summary_df[[i,"divisible_by_6"]] != "true" |
      !is.na(summary_df[[i,"frameshifts"]]) |
      !is.na(summary_df[[i,"nonsense_mutation"]]) |
      !is.na(summary_df[[i,"mutated_stop_codon"]]),
    summary_df$warning[[i]] <- "Warning", summary_df$warning[[i]] <- "PASS")
}
warning_samples_count <- (sum(summary_df[["warning"]] != "PASS", na.rm = FALSE)) - (failed_samples_count)

# Select and Rename wanted final columns
summary_df <- select(summary_df,
    sample,
    `strain`,
    `num_input_reads`,
    `num_aligned_reads`,
    `num_consensus_n`,
    `genome_completeness_percent`,
    `mean_sequencing_depth`,
    `median_sequencing_depth`,
    `total_variants`,
    `num_snps`,
    `num_deletions`,
    `num_deletion_sites`,
    `num_insertions`,
    `num_insertion_sites`,
    genome_length,
    divisible_by_6,
    frameshifts,
    nonsense_mutation,
    `mutated_stop_codon`,
    warning,
    `qc_status`,
    `run_status`,
    `run_summary`
  )
summary_df <- summary_df %>%
  rename(
    "Sample" = `sample`,
    "Strain" = `strain`,
    "# Input Reads" = `num_input_reads`,
    "# Aligned Reads" = `num_aligned_reads`,
    "# Consensus Ns" = `num_consensus_n`,
    "Genome Completeness (%)" = `genome_completeness_percent`,
    "Mean Depth" = `mean_sequencing_depth`,
    "Median Depth" = `median_sequencing_depth`,
    "Total Variants" = `total_variants`,
    "# SNPs" = `num_snps`,
    "# Deletions" = `num_deletions`,
    "# Deletion Sites" = `num_deletion_sites`,
    "# Insertions" = `num_insertions`,
    "# Insertion Sites" = `num_insertion_sites`,
    "Genome Length" = genome_length,
    "Divisible by 6" = divisible_by_6,
    "Frameshifts" = frameshifts,
    "Premature Stop Codon" = nonsense_mutation,
    "Mutated Stop Codon" = mutated_stop_codon,
    "QC Status" = `qc_status`,
    "Run Status" = `run_status`,
    "Run Summary" = `run_summary`
  )

# Set empty strains to NA for highlighting later
summary_df$Strain[is.na(summary_df$Strain)] <- "NA"

# Remove failed samples to be their own dataframe
if (length(failed_samples) > 0) {
  failed_df <- summary_df[summary_df$Sample %in% failed_samples, ]
  summary_df <- summary_df[!summary_df$Sample %in% failed_samples, ]
  failed_df <- select(failed_df,
    Sample,
    Strain,
    `# Input Reads`,
    `# Aligned Reads`,
    `Mean Depth`,
    `Median Depth`,
    `Genome Completeness (%)`
  )
}
```

<!-- ------------ -->
<!-- Report Start -->
<!-- ------------ -->
# Run Summary

## Row

```{r create-valueboxes, include=FALSE, message=FALSE, warning=FALSE}
# Create valueboxes for display
value_boxes <- list(
  reference_vbox = valueBox(strain, "Reference Strain", icon = "", color = "#343A40"),
  total_samples_vbox = valueBox(total_samples_count, "Total Samples", icon = "", color = "#D1D2D1"),
  passing_samples_vbox = valueBox(passing_samples_count, "Passing Samples", icon = "ion-android-checkmark-circle", color = "#18BC9C"),
  warning_samples_vbox = valueBox(warning_samples_count, "Samples with Warnings", icon = "ion-alert-circled", color = "#F39C12"),
  failed_samples_vbox = valueBox(failed_samples_count, "Failed Samples", icon = "ion-android-remove-circle", color = "#E74C3C")
)
```

```{r print-valueboxes, echo=FALSE, error=TRUE, results="asis"}
# Display valueboxes
for (box in value_boxes) {
  catHeader("", level = 3)
  print(box)
}
```

## Row

### **Table I**: Summary statistics for non-failing samples generated by the nf-core/viralrecon pipeline.

```{r interactive-table-display}
metrics_dt <- summary_df %>%
  datatable(
    rownames = FALSE,
    options = list(
      pageLength = 10,
      scrollX = FALSE,
      scrollY = "",
      columnDefs = list(
        list(className = 'dt-left', targets = "_all"),
        list(visible=FALSE, targets='warning')
      )
    )
  ) %>%
  # Row Highlighting
  formatStyle(
      'warning',
      target = 'row',
      backgroundColor = styleEqual(c("Pass", "Warning"),
      c('', '#F7BF65')
    )
  ) %>%
  # Cell Highlighting
  # Highlight cells where there are gene mutations
  formatStyle(
    columns = c("Mutated Stop Codon", "Premature Stop Codon", "Frameshifts"),
    target = "cell",
    fontWeight = styleEqual(
      unique(unlist(summary_df[c("Mutated Stop Codon", "Premature Stop Codon", "Frameshifts")])),
      ifelse(unique(unlist(summary_df[c("Mutated Stop Codon", "Premature Stop Codon", "Frameshifts")])) != "" , "bold", "")
    ),
    `text-decoration` = styleEqual(
      unique(unlist(summary_df[c("Mutated Stop Codon", "Premature Stop Codon", "Frameshifts")])),
      ifelse(unique(unlist(summary_df[c("Mutated Stop Codon", "Premature Stop Codon", "Frameshifts")])) != "" , "underline", "")
    )
  ) %>%
  # Highlight cells where the length is not divisible by 6
  formatStyle(
    'Divisible by 6',
    target = "cell",
    fontWeight = styleEqual(
      unique(summary_df$'Divisible by 6'),
      ifelse(unique(tolower(summary_df$'Divisible by 6')) != "true", "bold", "")
    ),
    `text-decoration` = styleEqual(
      unique(summary_df$'Divisible by 6'),
      ifelse(unique(tolower(summary_df$'Divisible by 6')) != "true", "underline", "")
    )
  ) %>%
  # Highlight cells where the nextclade strain doesn't match the reference strain
  formatStyle(
    'Strain',
    target = "cell",
    fontWeight = styleEqual(
      unique(summary_df$Strain),
      ifelse(unique(summary_df$'Strain') != strain, "bold", "")
    ),
    `text-decoration` = styleEqual(
      unique(summary_df$Strain),
      ifelse(unique(summary_df$'Strain') != strain, "underline", "")
    )
  ) %>%
  # Bars to show at a glance
  formatStyle(
    'Genome Completeness (%)',
    background = styleColorBar(range(0,100), 'lightblue', angle = -90),
    backgroundSize = '88% 88%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'left'
  )

metrics_dt
```

## Row {data-height=50}

<footer>
`r footer_str`
</footer>

<!-- --------------------- -->
<!-- Failed Sample Summary -->
<!-- --------------------- -->
```{r, failed-sample-page, child="subpage_failed.Rmd", params=params, eval=exists("failed_df")}
```

<!-- ---------------------------------------------------------- -->
<!-- Create Individual samples pages with the below code blocks -->
<!-- ---------------------------------------------------------- -->
```{r, render subpages, include=FALSE, echo=FALSE, error=TRUE}
# Code block to generate sample subpages - https://somtom.github.io/post/using-dynamically-rendered-r-markdown-childs-for-reports/
subpages = NULL
# Set knitr options to allow duplicate labels
options(knitr.duplicate.label = 'allow')
# Create temporary environment which we use for knitting subpages.RMD
subpage_env <- new.env()

# Create subpages
for ( sample_v in samples ) {

  # Filter data for the sample only
  if (sample_v %in% failed_samples) {
    sample_df <- failed_df %>%
      filter(Sample == sample_v)
  } else {
    sample_df <- summary_df %>%
      filter(Sample == sample_v)
  }

  # Assign data to subpage_env
  assign("sample_df", sample_df, subpage_env)
  assign("sample", sample_v, subpage_env)

  # Knit sample subpage using the subpage_env and add result to out vector
  subpages = c(subpages, knitr::knit_child('subpage_sample.Rmd', envir = subpage_env))
}
```

`r paste(knitr::knit_child(text = subpages), collapse = '')`

<!-- ---------------------- -->
<!-- Back to normal Rmd doc -->
<!-- ---------------------- -->

# Contact Information

<style>
/* Flip Box Container */
.flip-box {
  background-color: transparent;
  width: 700px;
  height: 400px;
  border: 1px solid #ccc;
  perspective: 1000px;
  margin: auto;
}

/* Flip Box Inner */
.flip-box-inner {
  position: relative;
  width: 100%;
  height: 100%;
  text-align: center;
  transition: transform 0.6s;
  transform-style: preserve-3d;
}

/* Flip Box on Hover */
.flip-box:hover .flip-box-inner {
  transform: rotateY(180deg);
}

/* Front & Back Side */
.flip-box-front, .flip-box-back {
  position: absolute;
  width: 100%;
  height: 100%;
  backface-visibility: hidden;
  display: flex;
  align-items: center;
  justify-content: center;
  border-radius: 10px;
}

/* Front Side */
.flip-box-front {
  background-color: #2C3E50;
  color: white !important;
}

/* Back Side */
.flip-box-back {
  background-color: #95A5A6;
  color: white !important;
  transform: rotateY(180deg);
}
</style>


```{r contact-info, warnings=FALSE, echo=FALSE}
if (any(nzchar(name), nzchar(phone), nzchar(email), nzchar(website))){
  back_content <- div(
    if (name != "") h2(name),
    if (email != "") p(strong("Email:"), email),
    if (phone != "") p(strong("Phone Number:"), phone),
    if (website != "") p("Website: ", a(href= paste0(website), website))
  )
} else {
  back_content <- div(
    h2("Contact Information Not Provided")
  )
}
flip_box( 
  front = div( 
    h2("Contact Info"), p("Click to reveal details"), icon("address-card", class = "fa-6x") 
  ), 
  back = back_content
)
```

<footer>
`r footer_str`
</footer>
