---
title: "MeaSeq: Measles Sequencing Run Report"
output: 
  flexdashboard::flex_dashboard:
    self_contained: false
    orientation: rows
    vertical_layout: scroll
    mathjax: NULL
    theme:
      version: 4
      bootswatch: flatly
    logo: phac.svg
knitr:
  opts_chunk:
    fig_embed: false
params:
  genotype: ""
  overall_qc: ""
  name: ""
  email: ""
  phone: ""
  website: ""
---

<style>

.navbar-logo img {
  position: relative;
  margin-left: 5px;
  height: 30px;
}
.navbar-fixed-top {
  margin-bottom: 20px;
}

body {
  padding-top: 60px;
}

footer {
  position: fixed;
  bottom: 0;
  width: 100%;
  background-color: #f1f1f1;
  text-align: center;
  padding: 0px;
  font-size: 12px;
  color: #333;
}
</style>

<!-- --------------------------------------- -->
<!-- Setup: libraries, parameters, functions -->
<!-- --------------------------------------- -->
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
# Load in libraries
suppressPackageStartupMessages(library(rmarkdown))
suppressPackageStartupMessages(library(flexdashboard))
suppressPackageStartupMessages(library(plotly))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(htmltools))
suppressPackageStartupMessages(library(shiny))
suppressPackageStartupMessages(library(shidashi))

# Assign imported parameters
genotype <- params$genotype
overall_qc <- params$overall_qc
name <- params$name
email <- params$email
phone <- params$phone
website <- params$website

# Create Header function
catHeader <- function(text = "", level = 3) {
  cat(
    paste0(
      "\n\n", 
      paste(rep("#", level), collapse = ""), 
      " ",
      text,
      "\n\n"
    )
  )
}

# Set report date
rep_date <- Sys.Date()

# Footer
footer_str <- paste(
  "Report generated using MeaSeq on",
  rep_date,
  "- © Government of Canada, Public Health Agency of Canada, National Microbiology Laboratory",
  sep = ' '
)

# Create function to give int values to qc statement for later colours
add_pass_value_int <- function(df, column) {
  # Split the given column, qc_status mostly, on ;
  split_vals <- strsplit(df[[column]], ';')

  # Define int values to go with statements
  key_phrases <- c('PASS', 'INCOMPLETE_GENOME')
  values <- c(1, 2)

  # Create a vector to store the results
  results <- integer(length(df[[column]]))

  # Add in the values
  for (i in 1:length(split_vals)) {
    # Check if any key phrase is contained in the split values
    matching_phrases <- intersect(split_vals[[i]], key_phrases)

    if (length(matching_phrases) > 0) {
      # If matching phrases are found, assign the corresponding integer value
      results[i] <- min(values[match(matching_phrases, key_phrases)])
    } else {
      results[i] <- 0  # If none of the key phrases are found, assign 0 for the other warnings
    }
  }

  return(results)
}

# Create a simplified and formatted version of sample to use to link to
create_sample_link <- function(sample) {
  sample <- tolower(sample)
  sample <- gsub("[^a-z0-9]+", "-", sample)
  sample <- gsub("^-+|-+$", "", sample)
  
  return(sample)
}

# Call a plotly plot so it can load later within for loops
htmltools::tagList(plot_ly(x = rnorm(10), type = "histogram"))

# Horizontal Line function for plotly
hline <- function(y = 0, color = "black") {
  list(
    type = "line",
    x0 = 0,
    x1 = 1,
    xref = "paper",
    y0 = y,
    y1 = y,
    line = list(color = color, dash = "dash")
  )
}
```

<!-- ------------------ -->
<!-- Setup: Sample Data -->
<!-- ------------------ -->
```{r interactive-table-setup}
# Read multiqc data
summary_df <- read_csv(overall_qc, locale = locale(encoding = "UTF-8"))
# Create a vector of sample names
samples <- summary_df$sample

# Add in 0's for specific NA value columns
summary_df <- summary_df %>%
  mutate_at(vars(`genome_completeness_percent`, `mean_sequencing_depth`, `median_sequencing_depth`), ~ ifelse(is.na(.), 0, .))

# Fill other empty columns with NA
summary_df[summary_df == ""] <- NA

# Get counts of passing, warning, and failing samples for display
total_samples_count <- nrow(summary_df)
failed_samples_count <- sum(summary_df[["genome_completeness_percent"]] < 50)
passing_samples_count <- total_samples_count - failed_samples_count
warning_samples_count <- ( sum(summary_df[["qc_status"]] != "PASS", na.rm = FALSE) - failed_samples_count )

# Add in colour code
summary_df$status_colour_code <- add_pass_value_int(summary_df, 'qc_status')

# Create vector of failed samples for the failed samples page
failed_samples <- summary_df %>%
  filter(`genome_completeness_percent` < 50) %>%
  pull(sample)

# Rename wanted final columns
#  Select them later for different tables
summary_df <- summary_df %>%
  rename(
    "Sample" = sample,
    "Genotype" = genotype,
    "Matched DSId" = matched_dsid,
    "# Input Reads" = num_input_reads,
    "# Aligned Reads" = num_aligned_reads,
    "# Consensus Ns" = num_consensus_n,
    "Genome Completeness (%)" = genome_completeness_percent,
    "Mean Depth" = mean_sequencing_depth,
    "Median Depth" = median_sequencing_depth,
    "Total Variants" = total_variants,
    "# SNPs" = num_snps,
    "# Deletions" = num_deletions,
    "# Deletion Sites" = num_deletion_sites,
    "# Insertions" = num_insertions,
    "# Insertion Sites" = num_insertion_sites,
    "Genome Length" = genome_length,
    "Divisible by 6" = divisible_by_6,
    "Frameshifts" = frameshifts,
    "Premature Stop Codon" = nonsense_mutation,
    "Mutated Stop Codon" = mutated_stop_codon,
    "QC Status" = qc_status,
    "Run Status" = run_status,
    "Run Summary" = run_summary
  )
```

<!-- ------------ -->
<!-- Report Start -->
<!-- ------------ -->
# Run Summary

## Row

```{r create-valueboxes, include=FALSE, message=FALSE, warning=FALSE}
# Create valueboxes for display
value_boxes <- list(
  reference_vbox = valueBox(genotype, "Reference Genotype", icon = "", color = "#343A40"),
  total_samples_vbox = valueBox(total_samples_count, "Total Samples", icon = "", color = "#D1D2D1"),
  passing_samples_vbox = valueBox(passing_samples_count, "Passing Samples", icon = "ion-android-checkmark-circle", color = "#18BC9C"),
  warning_samples_vbox = valueBox(warning_samples_count, "Samples with Warnings", icon = "ion-alert-circled", color = "#ffc773"),
  failed_samples_vbox = valueBox(failed_samples_count, "Failed Samples", icon = "ion-android-remove-circle", color = "#ff5147")
)
```

```{r print-valueboxes, echo=FALSE, error=TRUE, results="asis"}
# Display valueboxes
for (box in value_boxes) {
  catHeader("", level = 3)
  print(box)
}
```

## Row

### {-}

```{r interactive-table-display}
# Select the wanted summary metrics
metrics_df <- select(summary_df,
  Sample,
  Genotype,
  `Matched DSId`,
  `# Input Reads`,
  `# Aligned Reads`,
  `Mean Depth`,
  `Median Depth`,
  `Total Variants`,
  `Genome Completeness (%)`,
  `Genome Length`,
  `QC Status`,
  status_colour_code
)

# Add in a link for each sample
metrics_df$Sample <- sprintf(
  '<a href="MeaSeq_Report.html#%s" style="color: #000000;">%s</a>',
  create_sample_link(metrics_df$Sample),
  metrics_df$Sample
)

# Create table
metrics_dt <- metrics_df %>%
  datatable(
    rownames = FALSE,
    escape = FALSE,
    caption = htmltools::tags$caption(
      style = 'caption-side: top',
      htmltools::strong('Table I: '), 'Full Sample Summary'
    ),
    options = list(
      pageLength = 10,
      scrollX = FALSE,
      scrollY = "",
      columnDefs = list(
        list(className = 'dt-left', targets = "_all"),
        list(visible=FALSE, targets='status_colour_code')
      )
    )
  ) %>%
  # Row Highlighting
  #  1 - pass, 2 - fail, 3 - warning
  formatStyle(
      'status_colour_code',
      target = 'row',
      backgroundColor = styleEqual(c(1, 2, 0),
        c('', '#ff5147', '#ffc773')
    )
  ) %>%
  # Bars to show completeness at a glance
  formatStyle(
    'Genome Completeness (%)',
    background = styleColorBar(range(0,100), 'lightblue', angle = -90),
    backgroundSize = '88% 88%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'left'
  )

metrics_dt
```

## Row {data-height=400}

### {-}

```{r static-depth-report, fig.width=12, fig.height=3}
# Getting all the depth TSV files
files <- list.files("positional_depth", pattern = "\\.tsv$", full.names = TRUE)

# Use lapply to create list of dataframes adapted from the TSVs
df_list <- lapply(files, function(file) {
  sample <- tools::file_path_sans_ext(basename(file))
  df <- read_tsv(file, col_names = c("chrom", "position", "depth"), col_types = "cii")
  df$sample <- sample
  return(df)
})

# Combine all dfs
long_df <- bind_rows(df_list)

# Make sure position is numeric
long_df <- long_df %>% mutate(position = as.numeric(position))

# Summarize positions
#  Ymin and Ymax for drawing the stdev region
#  Ymin should not be below 0
summary_depth_df <- long_df %>%
  group_by(position) %>%
  summarise(
    mean = mean(depth, na.rm = TRUE),
    sd = sd(depth, na.rm = TRUE),
    median = median(depth, na.rm = TRUE),
    ymin = mean - sd,
    ymax = mean + sd,
    .groups = "drop"
  )
summary_depth_df$ymin[summary_depth_df$ymin < 0] <- 0

# Plot static image for now
max_pos <- max(summary_depth_df$position, na.rm = TRUE)
p <- ggplot(summary_depth_df, aes(x = position)) +
  geom_ribbon(aes(ymin = ymin, ymax = ymax), fill = "blue", alpha = 0.1) +
  geom_line(aes(y = median), color = "blue") +
  scale_x_continuous(
    breaks = seq(0, max_pos, by = 2000),
    minor_breaks = seq(0, max_pos, by = 500)
  ) +
  labs(title = "Median genomic read depth with deviation",
       x = "Genome position", y = "Median read depth") +
  theme_classic() +
  theme(
    panel.grid.minor.x = element_line(color = "#dadce0", linetype = "dotted"),
    panel.grid.major.x = element_line(color = "#dadce0", linetype = "dotted")
  )

p
```

> **Figure 1.** Median summarized genomic read depth with standard deviation. The median depth and standard deviation were calculated for each genomic position and plotted such that the median depth is highlighted plus and minus 1 standard deviation.

## Row {data-height=400}

### {-}

```{r static-depth-report-norm, fig.width=12, fig.height=3}
# Adapted from Molly's summary code we used for the paper
ndepth_raw <- read_csv("sample_positional_normalized_depth.csv")
ndepth <- ndepth_raw %>% rename(sample = 1)

transformed_data <- ndepth %>%
  pivot_longer(
    cols = -sample,
    names_to = "position",
    names_transform = list(position = as.integer),
    values_to = "norm_depth"
  )

# Summarize positions
#  Ymin and Ymax for drawing the stdev region
#  Ymin should never be below 0
#  Ymax capped to 2x max median to not explode chart
summary_depth_df <- transformed_data %>%
  group_by(position) %>%
  summarise(
    mean = mean(norm_depth),
    sd = sd(norm_depth),
    median = median(norm_depth),
    ymin = mean - sd,
    ymax = mean + sd,
    .groups = "drop"
  )
max_y_pos <- (2 * max(summary_depth_df$median))
summary_depth_df$ymin[summary_depth_df$ymin < 0] <- 0
summary_depth_df$ymax[summary_depth_df$ymax > max_y_pos] <- max_y_pos

# Plot static image, not sure how well an interactive one would be here
max_pos <- max(summary_depth_df$position)
p <- ggplot(summary_depth_df, aes(x = position)) +
  geom_ribbon(aes(ymin = ymin, ymax = ymax), fill = "blue", alpha = 0.1) +
  geom_line(aes(y = median), color = "blue") +
  scale_x_continuous(
    breaks = seq(0, max_pos, by = 2000),
    minor_breaks = seq(0, max_pos, by = 500)
  ) +
  labs(title = "Normalized variation of genomic read depth with deviation",
       x = "Genome position", y = "Proportional variation around mean depth") +
  geom_hline(yintercept = 1, linetype = "dashed", color = "black") +
  theme_classic() +
  # To better track where whole run positional depth is falling add in guide lines at minor axes
  theme(
    panel.grid.minor.x = element_line(color = "#dadce0", linetype = "dotted"),
    panel.grid.major.x = element_line(color = "#dadce0", linetype = "dotted")
  )

p
```

> **Figure 2.** Normalized variation of genomic read depth with standard deviation. Normalized variation of genomic read depth for each position in a sample was calculated using the positional depth divied by the mean. All positions were summarized by plotting the median read depth plus or minus 1 standard deviation surrounding the median with a minimum value of 0 and a maximum value of 2x the max median.

## Row {data-height=50}

<footer>
`r footer_str`
</footer>

<!-- ---------------------------------------------------------- -->
<!-- Create Individual samples pages with the below code blocks -->
<!-- ---------------------------------------------------------- -->
```{r, render subpages, include=FALSE, echo=FALSE, error=TRUE}
# Code block to generate sample subpages - https://somtom.github.io/post/using-dynamically-rendered-r-markdown-childs-for-reports/
subpages = NULL
# Set knitr options to allow duplicate labels
options(knitr.duplicate.label = 'allow')
# Create temporary environment which we use for knitting subpages.RMD
subpage_env <- new.env()

# Create subpages
for ( sample_v in samples ) {

  # Filter data for the sample only
  sample_df <- summary_df %>%
    filter(Sample == sample_v)

  # Assign data to subpage_env
  assign("sample_df", sample_df, subpage_env)
  assign("sample", sample_v, subpage_env)

  # Knit sample subpage using the subpage_env and add result to out vector
  subpages = c(subpages, knitr::knit_child('subpage_sample.Rmd', envir = subpage_env))
}
```

`r paste(knitr::knit_child(text = subpages), collapse = '')`

<!-- ---------------------- -->
<!-- Back to normal Rmd doc -->
<!-- ---------------------- -->

# Contact Information

<style>
/* Flip Box Container */
.flip-box {
  background-color: transparent;
  width: 700px;
  height: 400px;
  border: 1px solid #ccc;
  perspective: 1000px;
  margin: auto;
}

/* Flip Box Inner */
.flip-box-inner {
  position: relative;
  width: 100%;
  height: 100%;
  text-align: center;
  transition: transform 0.6s;
  transform-style: preserve-3d;
}

/* Flip Box on Hover */
.flip-box:hover .flip-box-inner {
  transform: rotateY(180deg);
}

/* Front & Back Side */
.flip-box-front, .flip-box-back {
  position: absolute;
  width: 100%;
  height: 100%;
  backface-visibility: hidden;
  display: flex;
  align-items: center;
  justify-content: center;
  border-radius: 10px;
}

/* Front Side */
.flip-box-front {
  background-color: #2C3E50;
  color: white !important;
}

/* Back Side */
.flip-box-back {
  background-color: #95A5A6;
  color: white !important;
  transform: rotateY(180deg);
}
</style>


```{r contact-info, warnings=FALSE, echo=FALSE}
if (any(nzchar(name), nzchar(phone), nzchar(email), nzchar(website))){
  back_content <- div(
    if (name != "") h2(name),
    if (email != "") p(strong("Email:"), email),
    if (phone != "") p(strong("Phone Number:"), phone),
    if (website != "") p("Website: ", a(href= paste0(website), website))
  )
} else {
  back_content <- div(
    h2("Contact Information Not Provided")
  )
}
flip_box( 
  front = div( 
    h2("Contact Info"), p("Click to reveal details"), icon("address-card", class = "fa-6x") 
  ), 
  back = back_content
)
```

<footer>
`r footer_str`
</footer>
