---
title: "MeaSeq: Measles Sequencing Run Report"
output: 
  flexdashboard::flex_dashboard:
    self_contained: true
    orientation: rows
    vertical_layout: scroll
    mathjax: NULL
    theme:
      version: 4
      bootswatch: flatly
knitr:
  opts_chunk:
    fig_embed: false
params:
  min_depth: 10
  version: ""
  revision: ""
  nf_version: ""
  overall_qc: ""
  variant_threshold: 0.75
  platform: ""
  name: ""
  email: ""
  phone: ""
  website: ""
---

```{css, echo=FALSE}
a, a:visited {
  color: #4A6A8A;
}

a:hover, a:focus {
  color: #2C3E50;
  text-decoration: underline;
}
```

<style>

.navbar-logo img {
  position: relative;
  margin-left: 5px;
  height: 30px;
}
.navbar-fixed-top {
  margin-bottom: 20px;
}

body {
  padding-top: 60px;
}

footer {
  position: fixed;
  bottom: 0;
  width: 100%;
  background-color: #f1f1f1;
  text-align: center;
  padding: 0px;
  font-size: 12px;
  color: #333;
}
</style>

<!-- Just to inject version into the navbar -->
```{r version_injection, echo=FALSE, results='asis'}
cat('
<script>
  document.addEventListener("DOMContentLoaded", function() {
    var navbar = document.querySelector(".navbar");
    if (navbar) {
      var versionDiv = document.createElement("div");
      versionDiv.textContent = "Pipeline Version ', params$version, '";
      versionDiv.style.position = "relative";
      versionDiv.style.right = "12px";
      versionDiv.style.color = "white";
      versionDiv.style.fontSize = "14px";
      versionDiv.style.fontWeight = "bold";
      versionDiv.style.zIndex = "1000";
      navbar.appendChild(versionDiv);
    }
  });
</script>
')
```

<!-- --------------------------------------- -->
<!-- Setup: libraries, parameters, functions -->
<!-- --------------------------------------- -->
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
# Load in libraries
suppressPackageStartupMessages(library(rmarkdown))
suppressPackageStartupMessages(library(flexdashboard))
suppressPackageStartupMessages(library(plotly))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(htmltools))
suppressPackageStartupMessages(library(shiny))
suppressPackageStartupMessages(library(shidashi))

# Assign imported parameters
overall_qc <- params$overall_qc
name <- params$name
email <- params$email
phone <- params$phone
website <- params$website

# Run contact sheet
run_contact <- any(nzchar(name), nzchar(phone), nzchar(email), nzchar(website))

# Create Header function
catHeader <- function(text = "", level = 3) {
  cat(
    paste0(
      "\n\n", 
      paste(rep("#", level), collapse = ""), 
      " ",
      text,
      "\n\n"
    )
  )
}

# Set report date
rep_date <- Sys.Date()

# Footer
footer_str <- paste(
  "Report generated using MeaSeq on",
  rep_date,
  "- © Government of Canada, Public Health Agency of Canada, National Microbiology Laboratory",
  sep = ' '
)

# Create function to give int values to qc statement for later colours
add_pass_value_int <- function(df, column) {
  # Split the given column, qc_status mostly, on ;
  split_vals <- strsplit(df[[column]], ';')

  # Define int values to go with statements
  key_phrases <- c('PASS', 'INCOMPLETE_GENOME')
  values <- c(1, 2)

  # Create a vector to store the results
  results <- integer(length(df[[column]]))

  # Add in the values
  for (i in 1:length(split_vals)) {
    # Check if any key phrase is contained in the split values
    matching_phrases <- intersect(split_vals[[i]], key_phrases)

    if (length(matching_phrases) > 0) {
      # If matching phrases are found, assign the corresponding integer value
      results[i] <- min(values[match(matching_phrases, key_phrases)])
    } else {
      results[i] <- 0  # If none of the key phrases are found, assign 0 for the other warnings
    }
  }

  return(results)
}

# Function to set HTML section ID based on alphanumeric sample name
make_sample_tab <- function(x) {
  x <- tolower(as.character(x))
  x <- gsub("[^a-z0-9]+", "-", x)
  x <- gsub("^-+|-+$", "", x)
  paste0("sample-", x)
}

# Call a plotly plot so it can load later within for loops
htmltools::tagList(plot_ly(x = rnorm(10), type = "histogram"))

# Horizontal Line function for plotly
hline <- function(y = 0, color = "black") {
  list(
    type = "line",
    x0 = 0,
    x1 = 1,
    xref = "paper",
    y0 = y,
    y1 = y,
    line = list(color = color, dash = "dash")
  )
}
```

<!-- ------------------ -->
<!-- Setup: Sample Data -->
<!-- ------------------ -->
```{r interactive-table-setup}
# Read multiqc data
summary_df <- read_csv(overall_qc, locale = locale(encoding = "UTF-8"))
# Create a vector of sample names
samples <- summary_df$sample

# Add in 0's for specific NA value columns
summary_df <- summary_df %>%
  mutate_at(vars(`genome_completeness_percent`, `mean_sequencing_depth`, `median_sequencing_depth`), ~ ifelse(is.na(.), 0, .))

# Fill other empty columns with NA
summary_df[summary_df == ""] <- NA

# Get counts of passing, warning, and failing samples for display
total_samples_count <- nrow(summary_df)
failed_samples_count <- sum( with(summary_df, genome_completeness_percent < 50 & qc_status != "PASS") )
passing_samples_count <- total_samples_count - failed_samples_count
warning_samples_count <- ( sum(summary_df[["qc_status"]] != "PASS", na.rm = FALSE) - failed_samples_count )

# Add in colour code
summary_df$status_colour_code <- add_pass_value_int(summary_df, 'qc_status')

# Create vector of failed samples for the failed samples page
failed_samples <- summary_df %>%
  filter(genome_completeness_percent < 50 & qc_status != "PASS") %>%
  pull(sample)

# Determine if we have DSIds to summarize
run_dsid <- any(!is.na(summary_df$matched_dsid) & nzchar(summary_df$matched_dsid))

# Based on platform set IUPAC or Masked sites column
iupac_final_col <- "# IUPACs"
iupac_initial_col <- "num_iupac"
if (params$platform == 'nanopore') {
  iupac_final_col <- "# Masked"
  iupac_initial_col <- "num_masked"
}

# Extract the most prevalent DSId for use later
extracted_dsids <- summary_df %>%
  filter(!matched_dsid %in% c("No Data", "Incomplete", "Ambiguous Base")) %>%
  count(matched_dsid, sort = TRUE)

prevalent_dsid <- case_when(
  nrow(extracted_dsids) == 0 ~ "No Data",
  sum(extracted_dsids$n == max(extracted_dsids$n)) > 1 ~ paste0(max(extracted_dsids$n)," DSIds"),
  TRUE ~ extracted_dsids$matched_dsid[1]
)

# Get the name of the DSId file used
dsid_file <- summary_df$dsid_file_used[1]

# Rename wanted final columns
#  Select them later for different tables
summary_df <- summary_df %>%
  rename(
    "Sample" = sample,
    "Genotype" = genotype,
    "Reference" = reference,
    "Matched DSId" = matched_dsid,
    "DSId File Used" = dsid_file_used,
    "# Input Reads" = num_input_reads,
    "# Aligned Reads" = num_aligned_reads,
    "# Consensus Ns" = num_consensus_n,
    "Genome Completeness (%)" = genome_completeness_percent,
    "Mean Depth" = mean_sequencing_depth,
    "Median Depth" = median_sequencing_depth,
    "Total Variants" = total_variants,
    "# SNPs" = num_snps,
    !!iupac_final_col := all_of(iupac_initial_col),
    "# Deletions" = num_deletions,
    "# Deletion Sites" = num_deletion_sites,
    "# Insertions" = num_insertions,
    "# Insertion Sites" = num_insertion_sites,
    "Genome Length" = genome_length,
    "Divisible by 6" = divisible_by_6,
    "Frameshifts" = frameshifts,
    "Premature Stop Codon" = nonsense_mutation,
    "Mutated Stop Codon" = mutated_stop_codon,
    "N450 Completeness" = N450_completeness,
    "N450 Mean Depth" = N450_mean_depth,
    "N450 Status" = N450_status,
    "QC Status" = qc_status,
    "Run Status" = run_status,
    "Run Summary" = run_summary
  )
```

```{r static-depth-report-setup}
# Getting all the depth TSV files
files <- list.files("positional_depth", pattern = "\\.tsv$", full.names = TRUE)

# Parse genomes and their corresponding samples
parsed_ids <- function(path) {
  base <- tools::file_path_sans_ext(basename(path))
  base <- str_remove(base, "_depth$")
  m <- str_match(base, "^([^_]+)_(.+)$")
  tibble(genome = m[, 2], sample =m[, 3])
}
# Use lapply to create list of dataframes adapted from the TSVs
df_list <- lapply(files, function(file) {
  ids <- parsed_ids(file)
  df <- read_tsv(file, col_names = c("chrom", "position", "depth"), col_types = "cii")
  df$sample <- ids$sample
  df$genome <- ids$genome
  return(df)
})

# Combine all dfs
long_df <- bind_rows(df_list)

# Make sure position is numeric
long_df <- long_df %>% mutate(position = as.numeric(position))

# Summarize positions
#  Ymin and Ymax for drawing the stdev region
#  Ymin should not be below 0
summary_depth_df <- long_df %>%
  group_by(genome, position) %>%
  summarise(
    mean = mean(depth, na.rm = TRUE),
    sd = sd(depth, na.rm = TRUE),
    median = median(depth, na.rm = TRUE),
    ymin = pmax(mean - sd, 0),
    ymax = mean + sd,
    .groups = "drop"
  )

# Gene Info based on final length
#  Currently only full genomes (~15894bp) and genomes with the 5' and 3' UTR cut (~15678bp) as references so
#  use that to determine location of gene boxes for some plots instead of needing to load in a GFF (at the moment)
gene_label <- c("N", "N450", "P/V/C", "M", "F", "H", "L")
gene_color <- c("#12f5f1ff", "#0da9a7ff", "#C52E19", "#6C8645", "#F98400", "#77be78ff", "#F2AD00")

# Genomes around 15678bp
gene_start_short <- c(1, 1128, 1700, 3331, 5351, 7164, 9127)
gene_end_short <- c(1578, 1578, 3233, 4338, 7003, 9017, 15678)
gene_label_loc_short <- c(789, 1353, 2466, 3834, 6177, 8090, 12402)

# Genomes around 15894bp
gene_start_long <- c(108, 1235, 1807, 3438, 5458, 7271, 9234)
gene_end_long <- c(1685, 1685, 3330, 4445, 7110, 9124, 15785)
gene_label_loc_long <- c(896, 1460, 2568, 3941, 6284, 8197, 12509)

# Gene DF for ggplot
gene_df_short <- tibble(
  label = gene_label,
  start = gene_start_short,
  end = gene_end_short,
  label_loc = gene_label_loc_short,
  color = gene_color
)
gene_df_long <- tibble(
  label = gene_label,
  start = gene_start_long,
  end = gene_end_long,
  label_loc = gene_label_loc_long,
  color = gene_color
)

# Plot static image
global_max_y <- max(summary_depth_df$median, na.rm = TRUE)
box_height <- 0.10 * global_max_y
global_min_y <- -box_height
make_depth_plot <- function(genome_id, summary_depth_df, panel_label) {
  df_g <- summary_depth_df %>% filter(genome == genome_id)
  genome_length <- max(df_g$position, na.rm = TRUE)

  # Set gene DF based on genome length
  gene_df <- if ( genome_length < 15800) {
    gene_df_short
  } else {
    gene_df_long
  }

  # Scale genes boxes
  box_ymin <- global_min_y
  box_ymax <- 0
  text_y <- (box_ymin + box_ymax) / 2

  # Add scales to Gene DF
  gene_df <- gene_df %>% mutate(
    box_ymin = box_ymin,
    box_ymax = box_ymax,
    text_y = text_y
  )

  # Plot figure
  p <- ggplot(df_g, aes(x = position)) +
    geom_ribbon(aes(ymin = ymin, ymax = ymax), fill = "blue", alpha = 0.25) +
    geom_line(aes(y = median), color = "blue") +
    scale_x_continuous(
      limits = c(0, genome_length),
      breaks = seq(0, genome_length, by = 2000),
      minor_breaks = seq(0, genome_length, by = 500)
    ) +
    scale_y_continuous(
      limits = c(box_ymin, NA)
    ) +
    labs(title = paste0("Panel ", panel_label, ": ", genome_id, " median genomic read depth with deviation"),
        x = "Genome position", y = "Median read depth") +
    theme_classic() +
    theme(
      panel.grid.minor.x = element_line(color = "#dadce0", linetype = "dotted"),
      panel.grid.major.x = element_line(color = "#dadce0", linetype = "dotted"),
      plot.margin = margin(5.5, 5.5, 5.5, 5.5)
    ) +
    geom_rect(
      data = gene_df,
      aes(xmin = start, xmax = end, ymin = box_ymin, ymax = box_ymax),
      fill = gene_df$color,
      color = "black",
      inherit.aes = FALSE,
      show.legend = FALSE
    ) +
    geom_text(
      data = gene_df,
      aes(x = label_loc, y = text_y, label = label),
      size = 2.5,
      inherit.aes = FALSE
    )
  return(p)
}

# Identify multiple genomes and add panel letter
genomes <- sort(unique(summary_depth_df$genome))
panel_letters <- LETTERS[seq_along(genomes)]

# Save plots
plots <- Map(function(genome_id, panel_letters) make_depth_plot(genome_id, summary_depth_df, panel_letters), genomes, panel_letters)
```

<!-- ------------ -->
<!-- Report Start -->
<!-- ------------ -->
Run Summary
=================================

## Row

```{r create-valueboxes, include=FALSE, message=FALSE, warning=FALSE}
# Create valueboxes for display
value_boxes <- list(
  total_samples_vbox = valueBox(total_samples_count, "Total Samples", icon = "", color = "#D1D2D1"),
  passing_samples_vbox = valueBox(passing_samples_count, "Passing Samples", icon = "ion-android-checkmark-circle", color = "#18BC9C"),
  warning_samples_vbox = valueBox(warning_samples_count, "Samples with Warnings", icon = "ion-alert-circled", color = "#ffc773"),
  failed_samples_vbox = valueBox(failed_samples_count, "Failed Samples", icon = "ion-android-remove-circle", color = "#ff5147")
)
```

```{r print-valueboxes, echo=FALSE, error=TRUE, results="asis"}
# Display valueboxes
for (box in value_boxes) {
  catHeader("", level = 3)
  print(box)
}
```

## Row

### {-}

```{r interactive-table-display}
# Select the wanted summary metrics
metrics_df <- select(summary_df,
  Sample,
  Genotype,
  Reference,
  `Matched DSId`,
  `DSId File Used`,
  `# Aligned Reads`,
  `Mean Depth`,
  `Median Depth`,
  `Total Variants`,
  !!iupac_final_col,
  `Genome Completeness (%)`,
  `Genome Length`,
  `N450 Status`,
  `QC Status`,
  status_colour_code
)

# Add in a link for each sample
metrics_df$Sample <- sprintf(
  '<a href="#%s" style="color: #000000;">%s</a>',
  make_sample_tab(metrics_df$Sample),
  metrics_df$Sample
)

# Create table
metrics_dt <- metrics_df %>%
  datatable(
    rownames = FALSE,
    escape = FALSE,
    caption = htmltools::tags$caption(
      style = 'caption-side: top',
      htmltools::strong('Table I: '), 'Full Sample Summary'
    ),
    options = list(
      pageLength = 10,
      order = list(list(7, 'desc'), list(3, 'desc')),
      scrollX = FALSE,
      scrollY = "",
      columnDefs = list(
        list(className = 'dt-left', targets = "_all"),
        list(visible=FALSE, targets='status_colour_code')
      )
    )
  ) %>%
  # Row Highlighting
  #  1 - pass, 2 - fail, 3 - warning
  formatStyle(
      'status_colour_code',
      target = 'row',
      backgroundColor = styleEqual(c(1, 2, 0),
        c('', '#ff5147', '#ffc773')
    )
  ) %>%
  # Bars to show completeness at a glance
  formatStyle(
    'Genome Completeness (%)',
    background = styleColorBar(range(0,100), 'lightblue', angle = -90),
    backgroundSize = '88% 88%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'left'
  )

metrics_dt
```

```{r static-depth-report-figs, fig.width=12, fig.height=3, out.width="100%", results='asis'}
if (length(genomes) > 1) {
  cat("<footer>")
  cat(footer_str)
  cat("</footer>")
  cat("Median Depth Plots\n")
  cat("=================================\n")
} else {
  catHeader("Row {data-height=400}", level = 2)
  catHeader("{-}", level = 3)
}

for (g in genomes) {
  print(plots[[g]])
}
```

> **Figure 1.** Median summarized genomic read depth with standard deviation. The median depth and standard deviation were calculated for each genomic position and plotted such that the median depth is highlighted plus and minus 1 standard deviation.

## Row {data-height=50}

<footer>
`r footer_str`
</footer>

<!-- ---------------------------------------------------------- -->
<!-- Create Individual samples pages with the below code blocks -->
<!-- ---------------------------------------------------------- -->
```{r, render subpages, include=FALSE, echo=FALSE, error=TRUE}
# Code block to generate sample subpages - https://somtom.github.io/post/using-dynamically-rendered-r-markdown-childs-for-reports/
subpages = NULL
# Set knitr options to allow duplicate labels
options(knitr.duplicate.label = 'allow')
# Create temporary environment which we use for knitting subpages.RMD
subpage_env <- new.env()

# Create subpages
for ( sample_v in samples ) {

  # Filter data for the sample only
  sample_df <- summary_df %>%
    filter(Sample == sample_v)

  # Assign data to subpage_env
  assign("sample_df", sample_df, subpage_env)
  assign("sample", sample_v, subpage_env)
  assign("sample_tab_id", make_sample_tab(sample_v), subpage_env)
  assign("params", params, subpage_env)
  assign("iupac_final_col", iupac_final_col, subpage_env)

  # Knit sample subpage using the subpage_env and add result to out vector
  subpages = c(subpages, knitr::knit_child('subpage_sample.Rmd', envir = subpage_env))
}
```

`r paste(knitr::knit_child(text = subpages), collapse = '')`

<!-- ---------------------- -->
<!-- Back to normal Rmd doc -->
<!-- ---------------------- -->

<!-- ---------------------------------------------------------- -->
<!-- DSId Page if used                                          -->
<!-- ---------------------------------------------------------- -->

```{r dsid_page, child="subpage_dsid.Rmd", params=params, eval=run_dsid}
```

Provenance
=================================

## Row

### {-}

```{r pipeline-name}
valueBox('phac-nml/measeq', "Pipeline Name")
```

### {-}

```{r pipeline-version}
valueBox(params$version, "Pipeline Version")
```

### {-}

```{r pipeline-rev}
valueBox(params$revision, "Pipeline Revision")
```

### {-}

```{r nfversion}
valueBox(params$nf_version, "Nextflow Version")
```

## Row

### {-}

```{r version-table-display}
# Create versions table for tracking
#  Yaml file name shouldn't change
version_yaml <- yaml::read_yaml("measeq_software_mqc_versions.yml")

# Track to build table df
process <- c()
software <- c()
versions <- c()

# Loop over data to make df
for (key_process in names(version_yaml)) {
  process_map <- version_yaml[[key_process]]

  # Workflow captured above
  if (key_process == "Workflow") next

  for (software_name in names(process_map)) {
    software_version <- process_map[[software_name]]

    # Add to lists
    process <- c(process, key_process)
    software <- c(software, software_name)
    versions <- c(versions, software_version)
  }
}

# Create df
df <- data.frame("Process Name" = process, "Software" = software, "Version" = versions)

# Create datatable
version_dt <- df %>%
  datatable(
    rownames = FALSE,
    escape = FALSE,
    caption = htmltools::tags$caption(
      style = 'caption-side: top',
      htmltools::strong('Table IIII: '), 'Software and Process Version Tracking'
    ),
    class = 'cell-border stripe',
    options = list(
      scrollX = FALSE,
      scrollY = "",
      columnDefs = list(
        list(className = 'dt-left', targets = "_all")
      )
    )
  )
version_dt
```

<!-- ---------------------------------------------------------- -->
<!-- Contact Page if used                                       -->
<!-- ---------------------------------------------------------- -->

```{r contact_page, child="subpage_contact.Rmd", params=params, eval=run_contact}
```
