<!-- Dynamically create individual sample rmd pages under the same datanav menu -->
`r paste0('# ', sample, '{data-navmenu="Sample Reports"}')`

## Row

### Sample Name
```{r, echo=FALSE, error=TRUE}
valueBox(sample, icon = "")
```

### Genome Coverage
```{r, echo=FALSE, error=TRUE}
completeness <- sample_df["Genome Completeness (%)"][1]
valueBox(completeness, icon = "")
```

### Sample Status
```{r, echo=FALSE, error=TRUE}
status_color = "#18BC9C" # Pass colour
if (sample %in% failed_samples) {
  status <- "Failed"
  status_color <- "#ff5147"
} else {
  status <- sample_df["QC Status"][1]
  if (status != "PASS") {
    status_color <- "#ffc773"
  }
}
valueBox(
  status,
  icon = "",
  color = status_color
)
```

## Row

### {-}
```{r, expanded_summary_table, echo=FALSE, error=TRUE}
# Select wanted cols to display
sample_df <- select(sample_df,
  Genotype,
  Reference,
  `Matched DSId`,
  `# Aligned Reads`,
  `Mean Depth`,
  `Median Depth`,
  `Genome Completeness (%)`,
  `Genome Length`,
  `Total Variants`,
  `# SNPs`,
  !!iupac_final_col,
  `# Deletions`,
  `# Deletion Sites`,
  `# Insertions`,
  `# Insertion Sites`,
  Frameshifts,
  `Premature Stop Codon`,
  `Mutated Stop Codon`,
  `QC Status`,
)

sddt <- sample_df %>%
  datatable(rownames = FALSE,
    caption = htmltools::tags$caption(
      style = 'caption-side: top',
      htmltools::strong('Table II: '), 'Expanded Sample Summary Table'
    ),
    options = list(
      dom = 't',
      paging = FALSE,
      scrollX = FALSE,
      scrollY = "",
      columnDefs = list(
        list(className = 'dt-left', targets = "_all")
      )
    )
  ) %>%
  formatStyle(
    'Genome Completeness (%)',
    background = styleColorBar(range(0,100), 'lightblue', angle = -90),
    backgroundSize = '88% 88%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'left'
  )

sddt
```

## Row

### {-}
```{r, variant_plot, echo=FALSE, error=TRUE}
# Initialize gene boxes for plots based on genome length
if (sample_df$'Genome Length' < 15800 ) {
  gene_info <- gene_df_short
} else {
  gene_info <- gene_df_long
}
gene_shapes_normal <- lapply(seq_along(gene_info$start), function(i) {
  list(
    type = "rect",
    x0 = gene_info$start[i], x1 = gene_info$end[i],
    y0 = -3.5, y1 = -0.5,
    fillcolor = gene_info$color[i],
    line = list(color = "black"),
    opacity = 0.8,
    layer = "below"
  )
})
gene_shapes_log <- lapply(seq_along(gene_info$start), function(i) {
  list(
    type = "rect",
    x0 = gene_info$start[i], x1 = gene_info$end[i],
    y0 = 0.7, y1 = 0.9,
    fillcolor = gene_info$color[i],
    line = list(color = "black"),
    opacity = 0.8,
    layer = "below"
  )
})

#### Variants Plot ####
# Create table from bam info CSV showing variants and subvariants
var_f <- paste0("variation/", sample, "_variation.csv")

# Default data if it exists or doesn't
if (file.exists(var_f)) {
  variants <- read.table(var_f, header = TRUE, sep = ",")
} else {
  variants <- data.frame(matrix(ncol = 12, nrow = 1))
  colv <- c("position", "ref_base", "percentage_nonref", "variant_type", "total_reads", "A", "T", "C", "G", "del", "ins", "distribution")
  colnames(variants) <- colv
}

# Text info for hover
text_info <- paste(
    'Position: ', variants[['position']], '<br>',
    'Reference: ', variants[['ref_base']], '<br>',
    'Total Reads: ', variants[['total_reads']], '<br>',
    '  A: ', variants[['A']], '<br>',
    '  T: ', variants[['T']], '<br>',
    '  G: ', variants[['G']], '<br>',
    '  C: ', variants[['C']], '<br>',
    '  Del:', variants[['del']], '<br>',
    '  Ins:', variants[['ins']], '<br>',
    'Insertion Distribution: ', variants[['distribution']], '<br>',
    sep = ''
)
variants$text_info <- text_info

# Plot variation table
#  Also customdata instead of text due to https://github.com/plotly/plotly.R/issues/1859
p <- plot_ly(variants) %>%
  add_trace(x = variants[['position']], y = variants[['percentage_nonref']],
    type = 'scatter', mode = 'markers',
    color = ~factor(variant_type, levels = c("A SNP", "T SNP", "G SNP", "C SNP", "Ref", "Mixed", "Deletion", "Insertion")),
    marker = list(sizemode = 'diameter', size = 15.0, opacity = 0.7),
    customdata = ~text_info,
    hovertemplate = paste(
        '<b>Non-Reference Percentage</b>: %{y:.2f}<br>',
        '%{customdata}<extra></extra>',
        sep=''
    )
  )

#### Add in the Gene Annotations ####
final_thresh <- (params$variant_threshold * 100)
final_shapes <- c(gene_shapes_normal, list(hline(final_thresh)))

# Labels
p <- p %>% add_text(
  x = gene_info$label_loc,
  y = rep(-6.5, length(gene_info$label_loc)),
  text = gene_info$label,
  textfont = list(size = 14, family = "Arial Black"),
  showlegend = FALSE
)

# Add final layout to plot
p <- p %>% layout(
  shapes = final_shapes,
  xaxis = list(title = "Position", tickformat = ".0f"),
  yaxis = list(title = "Percent Non Reference Base", tickformat = '2.%')
)

p
```

> **Figure 2.** Subconsensus Variant Read Counts. The sample pileup was analyzed to find positions in the genome where the non-reference allele percentage was >=15%.
> All positions with >=8 reads (slightly less than the default minimum depth), a positional read base quality of >=0 and a map quality >=0 were included in the analysis.
> Mixed sites in the figure that are not in the variant table are variants that have failed processing and contain the reference bases in the final consensus
> if they are above the set minimum depth. The dashed line represents the threshold at which a variant is called if it passes all other filters.

## Row

### {-}
```{r, variant_table, echo=FALSE, error=TRUE}
#### Variants Table ####
# Get variants tsv data
var_f <- paste0("variant_tsv/", sample, ".consensus.tsv")
var_c_names <- c("pos", "ref", "alt", "qual", "depth", "vaf", "tag")

# Get depth data to confirm any low-depth variants
depth_f <- paste0("positional_depth/", sample_df$Reference, "_", sample, "_depth.tsv")
if (file.exists(depth_f) && file.info(depth_f)$size > 0) {
  coverage <- read.table(depth_f, sep = "\t", header = FALSE)
  colnames(coverage) <- c("chrom", "position", "depth")
} else {
  coverage <- data.frame(matrix(ncol = 3, nrow = 1))
  colv <- c("chrom", "position", "depth")
  colnames(coverage) <- colv
}

# Create table
if (file.exists(var_f)) {
  # Col classes incase there is few variants and they are all Ts (odd issue but yeah)
  #  Which then becomes 'true' which is not ideal
  col_classes <- c(NA, "integer", "character", "character", rep(NA, 4))
  variants <- read.table(var_f, header = TRUE, sep = "\t", colClasses = col_classes)
  variants <- select(variants, var_c_names)
  # Filter out failed variants
  variants <- variants[variants$tag != "Fail", ]
  # Filter out low-depth masked variants
  variants <- variants %>%
    left_join(coverage, by = c("pos" = "position")) %>%
    filter(depth.y >= params$min_depth) %>%
    select(-depth.y, -chrom)
  # Create table
  variants <- variants %>%
    rename(
      "Pos" = pos,
      "Ref" = ref,
      "Alt" = alt,
      "Qual" = qual,
      "Depth" = depth.x,
      "Alt Allele Fraction" = vaf,
      "Type" = tag
    )
  ddt <- variants %>%
    datatable(rownames = FALSE,
      caption = htmltools::tags$caption(
        style = 'caption-side: top',
        htmltools::strong('Table III: '), 'Consensus Variant Summary Table. VCF-based summary table of the variants found in the final consensus sequence. The "Type" column displays more information on the variant call such as if the site was masked (Nanopore) or an IUPAC base (Illumina).'
      ),
      options = list(
        pageLength = 10,
        scrollX = FALSE,
        scrollY = "",
        columnDefs = list(
          list(className = 'dt-left', targets = "_all")
        )
      )
    ) %>%
    formatRound(columns = c('Qual'), digits = 2) %>%
    formatRound(columns = c('Alt Allele Fraction'), digits = 4) %>%
    # Bars to show at a glance
    formatStyle(
      'Alt Allele Fraction',
      background = styleColorBar(range(0,1), 'lightblue', angle = -90),
      backgroundSize = '88% 88%',
      backgroundRepeat = 'no-repeat',
      backgroundPosition = 'left'
    )
} else {
  variants <- data.frame(matrix(ncol = 7, nrow = 1))
  colnames(variants) <- var_c_names
  ddt <- variants %>%
    datatable(rownames = FALSE)
}

ddt
```

## Row

### {-}
```{r, depth_plot, echo=FALSE, error=TRUE}
#### Depth Plots ####
# Data already gotten in last step
# Set depths of 0 to 1 to allow continuous visualization with log plot
coverage$depth[coverage$depth == 0] <- 1

# Create the background shapes needed
background_shapes <- list(
  list(
    type = "rect",
    fillcolor = "red",
    opacity = 0.2,
    x0 = min(coverage$position),
    x1 = max(coverage$position),
    y0 = 1,
    y1 = 10
  ),
  list(
    type = "rect",
    fillcolor = "yellow",
    opacity = 0.2,
    x0 = min(coverage$position),
    x1 = max(coverage$position),
    y0 = 10,
    y1 = 30
  ),
  list(
    type = "rect",
    fillcolor = "green",
    opacity = 0.2,
    x0 = min(coverage$position),
    x1 = max(coverage$position),
    y0 = 30,
    y1 = max(coverage$depth)
  )
)

final_shapes <- c(background_shapes, gene_shapes_log)

# Create plot
depthp <- plot_ly(x = ~coverage$position, y = ~coverage$depth, type = "scatter", mode = "lines", line = list(color = "black")) %>%
  layout(
    yaxis = list(
    type = "log",
    title = "Sequencing Depth (Log Scale)",
    tickvals = c(10, 30, 100, 1000),
    ticktext = c(10, 30, 100, 1000)
  ),
  xaxis = list(
    title = "Genomic Position (nt)",
    tickformat = ".0f"
  ),
  shapes = final_shapes
)

# Add labels
depthp <- depthp %>% add_text(
  x = gene_info$label_loc,
  y = rep(0.55, length(gene_info$label_loc)),
  text = gene_info$label,
  inherit = FALSE,
  type = "scatter",
  mode = "text",
  textfont = list(size = 14, family = "Arial Black"),
  showlegend = FALSE
) 

depthp
```
> **Figure 3.** Positional genomic depth. The Y-axis is log-scaled so positions with no genomic depth have been set to 1 to allow a continuous visualization. Zones have been marked using a colour scheme where: 0-10 is red corresponding to masked sites, 10-30 is yellow corresponding to sites where there is a basecall but limited depth, and 30+ is green for positions with acceptable or better depth.

## Row

### {-}
```{r, qual_plot, echo=FALSE, error=TRUE}
#### Quality Plots ####
# Data
quality <- read.table(paste0("positional_baseq/", sample, ".baseq.stats.tsv"), sep = "\t", header = TRUE)
custom_tickvals <- c(0,10,20,30)

# Create the background shapes needed
background_shapes <- list(
  list(
    type = "rect",
    fillcolor = "red",
    opacity = 0.2,
    x0 = min(quality$pos),
    x1 = max(quality$pos),
    y0 = 0,
    y1 = 20
  ),
  list(
    type = "rect",
    fillcolor = "yellow",
    opacity = 0.2,
    x0 = min(quality$pos),
    x1 = max(quality$pos),
    y0 = 20,
    y1 = 30
  ),
  list(
    type = "rect",
    fillcolor = "green",
    opacity = 0.2,
    x0 = min(quality$pos),
    x1 = max(quality$pos),
    y0 = 30,
    y1 = 60
  )
)

final_shapes <- c(background_shapes, gene_shapes_normal)

# Plot
qualityp <- plot_ly(quality ,x = ~quality$'pos', y = ~quality$'rms_baseq', type = "scatter", mode = "lines", line = list(color = "black")) %>%
  layout(
    yaxis = list(
      tickvals = custom_tickvals,
      ticktext = c("0", "10", "20", "30"),
      title = "RMS Base Quality Score"),
    xaxis = list(
      title = "Genomic Position (nt)",
      tickformat = ".0f"
    ),
    shapes = final_shapes
  )

# Add labels
qualityp <- qualityp %>% add_text(
  x = gene_info$label_loc,
  y = rep(-6.5, length(gene_info$label_loc)),
  text = gene_info$label,
  inherit = FALSE,
  type = "scatter",
  mode = "text",
  textfont = list(size = 14, family = "Arial Black"),
  showlegend = FALSE
)

qualityp
```
> **Figure 4.** Average root-mean-square read base quality score per genomic position

## Row {data-height=50}

<footer>
`r footer_str`
</footer>
