<!-- Dynamically create individual sample rmd pages under the same datanav menu -->
`r paste0('# ', sample, '{data-navmenu="Sample Reports"}')`

Row
------------------------------

### Sample Name
```{r, echo=FALSE, error=TRUE}
valueBox(sample, icon = "")
```

### Genome Coverage
```{r, echo=FALSE, error=TRUE}
completeness <- sample_df["Genome Completeness (%)"][1]
valueBox(completeness, icon = "")
```

### Sample Status
```{r, echo=FALSE, error=TRUE}
status_color = "#18BC9C"
if (sample %in% failed_samples) {
  status <- "Failed"
  status_color <- "#ff5147"
} else {
  status <- sample_df["QC Status"][1]
  if (status != "PASS") {
    status_color <- "#ffc773"
  }
}
valueBox(
  status,
  icon = "",
  color = status_color
)
```

Row
------------------------------

### {-}
```{r, expanded_summary_table, echo=FALSE, error=TRUE}
# Select wanted cols to display
sample_df <- select(sample_df,
  Genotype,
  `Matched DSId`,
  `# Aligned Reads`,
  `Mean Depth`,
  `Median Depth`,
  `Genome Completeness (%)`,
  `Genome Length`,
  `Total Variants`,
  `# SNPs`,
  `# Deletions`,
  `# Deletion Sites`,
  `# Insertions`,
  `# Insertion Sites`,
  Frameshifts,
  `Premature Stop Codon`,
  `Mutated Stop Codon`,
  `QC Status`,
)

sddt <- sample_df %>%
  datatable(rownames = FALSE,
    caption = htmltools::tags$caption(
      style = 'caption-side: top',
      htmltools::strong('Table II: '), 'Expanded Sample Summary Table'
    ),
    options = list(
      dom = 't',
      paging = FALSE,
      scrollX = FALSE,
      scrollY = "",
      columnDefs = list(
        list(className = 'dt-left', targets = "_all")
      )
    )
  ) %>%
  formatStyle(
    'Genome Completeness (%)',
    background = styleColorBar(range(0,100), 'lightblue', angle = -90),
    backgroundSize = '88% 88%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'left'
  )

sddt
```

Row
------------------------------

### {-}
```{r, variant_plot, echo=FALSE, error=TRUE}
#### Variants Plot ####
# Create table from bam info CSV showing variants and subvariants
var_f <- paste0("variation/", sample, "_variation.csv")

# Default data if it exists or doesn't
if (file.exists(var_f)) {
  variants <- read.table(var_f, header = TRUE, sep = ",")
} else {
  variants <- data.frame(matrix(ncol = 12, nrow = 1))
  colv <- c("position", "ref_base", "percentage_nonref", "variant_type", "total_reads", "A", "T", "C", "G", "del", "ins", "distribution")
  colnames(variants) <- colv
}

# Text info for hover
text_info <- paste(
    'Position: ', variants[['position']], '<br>',
    'Reference: ', variants[['ref_base']], '<br>',
    'Total Reads: ', variants[['total_reads']], '<br>',
    '  A: ', variants[['A']], '<br>',
    '  T: ', variants[['T']], '<br>',
    '  G: ', variants[['G']], '<br>',
    '  C: ', variants[['C']], '<br>',
    '  Del:', variants[['del']], '<br>',
    '  Ins:', variants[['ins']], '<br>',
    'Insertion Distribution: ', variants[['distribution']], '<br>',
    sep = ''
)
variants$text_info <- text_info

# Plot variation table
#  Also customdata instead of text due to https://github.com/plotly/plotly.R/issues/1859
p <- plot_ly(variants) %>%
  add_trace(x = variants[['position']], y = variants[['percentage_nonref']],
    type = 'scatter', mode = 'markers',
    color = ~factor(variant_type, levels = c("A SNP", "T SNP", "G SNP", "C SNP", "Ref", "Mixed", "Deletion", "Insertion")),
    marker = list(sizemode = 'diameter', size = 15.0, opacity = 0.7),
    customdata = ~text_info,
    hovertemplate = paste(
        '<b>Percentage</b>: %{y:.2f}<br>',
        '%{customdata}<extra></extra>',
        sep=''
    )
  )

# Add final layout to plot
p <- p %>%
  layout(
    xaxis = list (title = "Position", tickformat = ".0f"),
    yaxis = list (title = "Percent Non Reference Base", tickformat = '2.%'),
    shapes = hline(75)
  )

p
```

> **Figure 2.** Subconsensus Variant Read Counts. The sample pileup was analyzed to find positions in the genome where the non-reference allele percentage was >=15%. All positions with >=10 reads, a positional read base quality of >=10 and a map quality >=30 were included in the analysis.

## Row

### {-}
```{r, variant_table, echo=FALSE, error=TRUE}
#### Variants Table ####
# Get data
var_f <- paste0("variant_tsv/", sample, ".consensus.tsv")
var_c_names <- c("pos", "ref", "alt", "qual", "depth", "vaf", "tag")

# Create table
if (file.exists(var_f)) {
  variants <- read.table(var_f, header = TRUE, sep = "\t")
  variants <- select(variants, var_c_names)
  variants <- variants %>%
    rename(
      "Pos" = pos,
      "Ref" = ref,
      "Alt" = alt,
      "Qual" = qual,
      "Depth" = depth,
      "Variant Allele Fraction" = vaf,
      "Type" = tag
    )
  ddt <- variants %>%
    datatable(rownames = FALSE,
      caption = htmltools::tags$caption(
        style = 'caption-side: top',
        htmltools::strong('Table III: '), 'Consensus Variants Summary'
      ),
      options = list(
        pageLength = 10,
        scrollX = FALSE,
        scrollY = "",
        columnDefs = list(
          list(className = 'dt-left', targets = "_all")
        )
      )
    ) %>%
    formatRound(columns = c('Qual'), digits = 2) %>%
    formatRound(columns = c('Variant Allele Fraction'), digits = 4) %>%
    # Bars to show at a glance
    formatStyle(
      'Variant Allele Fraction',
      background = styleColorBar(range(0,1), 'lightblue', angle = -90),
      backgroundSize = '88% 88%',
      backgroundRepeat = 'no-repeat',
      backgroundPosition = 'left'
    )
} else {
  variants <- data.frame(matrix(ncol = 7, nrow = 1))
  colnames(variants) <- var_c_names
  ddt <- variants %>%
    datatable(rownames = FALSE)
}

ddt
```

Row
------------------------------

### {-}
```{r, depth_plot, echo=FALSE, error=TRUE}
#### Depth Plots ####
# Get data
coverage <- read.table(paste0("positional_depth/", sample, "_depth.tsv"), sep = "\t", header = FALSE)
colnames(coverage) <- c("chrom", "position", "depth")

# Set depths of 0 to 1 to allow continuous visualization with log plot
coverage$depth[coverage$depth == 0] <- 1
depthp <- plot_ly(x = ~coverage$position, y = ~coverage$depth, type = "scatter", mode = "lines", line = list(color = "black")) %>%
  layout(
    yaxis = list(
    type = "log",
    title = "Sequencing Depth (Log Scale)",
    tickvals = c(10, 30, 100, 1000),
    ticktext = c(10, 30, 100, 1000)
  ),
  xaxis = list(
    title = "Genomic Position (nt)",
    tickformat = ".0f"
  ),
  shapes = list(
    list(
      type = "rect",
      fillcolor = "red",
      opacity = 0.2,
      x0 = min(coverage$position),
      x1 = max(coverage$position),
      y0 = 1,
      y1 = 10
    ),
    list(
      type = "rect",
      fillcolor = "yellow",
      opacity = 0.2,
      x0 = min(coverage$position),
      x1 = max(coverage$position),
      y0 = 10,
      y1 = 30
    ),
    list(
      type = "rect",
      fillcolor = "green",
      opacity = 0.2,
      x0 = min(coverage$position),
      x1 = max(coverage$position),
      y0 = 30,
      y1 = max(coverage$depth)
    )
  )
)
depthp
```
> **Figure 3.** Positional genomic depth. The Y-axis is log-scaled so positions with no genomic depth have been set to 1 to allow a continuous visualization. Zones have been marked using a colour scheme where: 0-10 is red corresponding to masked sites, 10-30 is yellow corresponding to sites where there is a basecall but limited depth, and 30+ is green for positions with acceptable or better depth.

Row
------------------------------

### {-}
```{r, qual_plot, echo=FALSE, error=TRUE}
#### Quality Plots ####
# Data
quality <- read.table(paste0("positional_baseq/", sample, ".baseq.stats.tsv"), sep = "\t", header = TRUE)
custom_tickvals <- c(0,10,20,30)

# Plot
qualityp <- plot_ly(quality ,x = ~quality$'pos', y = ~quality$'rms_baseq', type = "scatter", mode = "lines", line = list(color = "black")) %>%
  layout(
    yaxis = list(
      tickvals = custom_tickvals,
      ticktext = c("0", "10", "20", "30"),
      title = "RMS Base Quality Score"),
    xaxis = list(
      title = "Genomic Position (nt)",
      tickformat = ".0f"
    ),
    shapes = list(
      list(
        type = "rect",
        fillcolor = "red",
        opacity = 0.2,
        x0 = min(quality$pos),
        x1 = max(quality$pos),
        y0 = 0,
        y1 = 20
      ),
      list(
        type = "rect",
        fillcolor = "yellow",
        opacity = 0.2,
        x0 = min(quality$pos),
        x1 = max(quality$pos),
        y0 = 20,
        y1 = 30
      ),
      list(
        type = "rect",
        fillcolor = "green",
        opacity = 0.2,
        x0 = min(quality$pos),
        x1 = max(quality$pos),
        y0 = 30,
        y1 = 60
      )
    )
  )
qualityp
```
> **Figure 4.** Average root-mean-square read base quality score per genomic position

Row
------------------------------

### {-}
```{r, n_content_plot, echo=FALSE, error=TRUE}
#### N Content Plots ####
n_content <- read.table(paste0("positional_n_depth/", sample, ".per_base_n.tsv"), sep = "\t", header = TRUE)
coverage <- read.table(paste0("positional_depth/", sample, "_depth.tsv"), sep = "\t", header = FALSE)
colnames(coverage) <- c("chrom", "position", "depth")

n_cov <- merge(n_content, coverage, by = "position")
n_cov$n_cov <- ((n_cov$N_density)/n_cov$depth)*100
n_cov$n_cov <- ifelse(is.nan(n_cov$n_cov), 0, n_cov$n_cov)
Q1 <- quantile(n_cov$n_cov, 0.25)
Q3 <- quantile(n_cov$n_cov, 0.75)
IQR <- Q3 - Q1
IQR <- unname(IQR)
outliers <- n_cov$n_cov[n_cov$n_cov < Q1 - 1.5 * IQR | n_cov$n_cov > Q3 + 1.5 * IQR ]
dr <- n_cov[(n_cov$n_cov %in% outliers), ]
# if (IQR == 0 | is.na(IQR) |is.nan(IQR)) {
#   cat("The figure for this sample couldn't be processed. The per-base N density across the sample is within the interquartile range for this sample with no outliers.\n\n")
# } else {
  nplot <- plot_ly(n_cov, x = n_cov$position, y = n_cov$n_cov, type = "scatter", mode = "lines", line = list(color = "#000000")) %>%
    layout(
      xaxis = list(
        title = "Genomic Position (nt)",
        tickformat = ".0f"
      ),
      yaxis = list(
        title = "Percentage (%) of N's"
      )
    )
  nplot
# }
```
> **Figure 5.** The percentage of base calls at each position for which an N was called along the genome. The displayed positions are those identified as outliers by using quartile ranges.

## Row {data-height=50}

<footer>
`r footer_str`
</footer>
